---
title: "Lesson 04: Probability"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
	base.dir <- lp.w
} else if (file.exists(lp.h)) {
	base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

```


```{r ex_3_util, echo = FALSE}

```

## 1.) Shoppers

+ __Dataset__: _shoppers.csv_

```{r ex_1_data, echo = FALSE}

shoppers <- loadDataFile("shoppers.csv")

glimpse(shoppers)

```

Assume the fifty shoppers exit the store individually in random order.

### a.)

> If one shopper is picked at random, what is the probability of picking a shopper who spent
$40 or more dollars? What is the probability of picking a shopper who spent less than $10?

```{r ex_1_a, echo = T}

n <- length(shoppers$Spending)

gte_40 <- sum(shoppers$Spending >= 40)
gte_40 / n

lt_10 <- sum(shoppers$Spending < 10)
lt_10 / n


```

### b.)

> If two shoppers are picked at random, what is the probability the pair will include a shopper
who spent $40 or more dollars and one who spent less than $10?

```{r ex_1_b, echo = T}

gte_40 * lt_10 / ( n * ( n - 1 ) / 2 )

```

_Hint : For parts c) and d) it will be necessary to assume sampling without replacement._

### c.) 

> If two shoppers are picked at random, what is the probability the pair will include two
shoppers who spent no less than $10 and no more than $40?

```{r}

m <- sum( (shoppers$Spending >= 10) & (shoppers$Spending <= 40))

m * (m - 1) / ( n * (n - 1) )

```

### d.)

> If four shoppers are picked at random, what is the probability one shopper will have spent
less than $10, one shopper will have spent $40 or more dollars and two shoppers will have
spent no less than $10 and no more than $40?

```{r}

successful_combinations <- gte_40 * lt_10 * m * (m - 1) / 2
total_combinations <- dim( combn( n, 4 ) )[ 2 ]

successful_combinations / total_combinations


```

### e.)

> If we know a randomly picked shopper has spent more than $30, what is the probability that
shopper has spent more than $40?

```{r}

g1 <- sum(shoppers$Spending > 30)
g2 <- sum(shoppers$Spending > 40)

g2 / g1

```

## 2.) R Sampling

Use R to answer these questions

### a.) 

+ Draw 100 samples with replacement of size 22 from the 365 integers (i.e. 1,2,.,365). 
+ Count the number of samples in which one or more of the numbers sampled is duplicated.
+ Divide by 100 to estimate the probability of such duplication occurring. 
+ (If 22 people are selected at random, what is the probability of two or more matching birthdays?)

```{r}

integers <- seq( from = 1, to = 365, by = 1 )

set.seed(1234) # fix the sampling as to reproduce

samples <- replicate( 100, sample(integers, 22, replace = TRUE))

dupes <- apply(samples, 2, function(x) {
	length(x) != length(unique(x))
})

sum(dupes) / 100

```

### b.)

_Suppose that 60% of marbles in a bag are black and 40% are white._

+ Generate a random sample of size 20 with replacement using uniform random numbers. 
+ For the numbers in each sample, if a random number is 0.6 or less, code it as a 1. If it is not 0.6 or less code it a zero.
+ Add the twenty coded numbers.

```{r}

population_size <- 50

gen_marbles <- function() { 
	sample <- sample(runif(population_size), size = 20, replace = T)
    coded <- ifelse( sample <= .6, 1, 0 )

    sum(coded)
}

num_results <- function(trials, value = 11) { 
    simulations <- replicate(trials, gen_marbles() )

	hist(simulations,
		breaks = pretty(simulations),
		col = "red",
        main = paste("Expected Value:", sum(simulations) / trials))

	sum(simulations >= value)
}

```

Do this 50 times and calculate the proportion of times the sum is 11 or greater. 

```{r}

    runs <- 50
    results <- num_results(runs)

    results / runs # probability
    
```

What have you estimated? Expand the number of trials to 10,000.

```{r}

    runs <- 10000
    results <- num_results(runs)

    results / runs # probability

```

The exact binomial estimated probability is 0.755 and the expectation is 12.
