---
title: "Lesson 03: Descriptive Statistics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
	base.dir <- lp.w
} else if (file.exists(lp.h)) {
	base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

```


```{r ex_3_util, echo = FALSE}

getMode <- function(values) {
	tbl.vals <- data.frame(table(values))
	tbl.vals$values <- as.numeric(levels(tbl.vals$values))[tbl.vals$values]
	mode <- tbl.vals[which.max(tbl.vals$Freq),]

	names(mode)[1] <- "Value"

	mode
}

getSummary <- function(values) {
	mode <- getMode(values)

	data.table(
		Statistic = c("Mean",
					"Median",
					"Min",
					"Max",
                    "Range",
					"Mode",
					"ModeFreq",
	                "StdDev",
					"Variance",
					"Q1",
					"Q3",
                    "P10"),

	    Value = c(mean(values, na.rm = T),
			  median(values, na.rm = T),
			  min(values, na.rm = T),
			  max(values, na.rm = T),
              max(values, na.rm = T) - min(values, na.rm = T),
			  ifelse(mode$Freq != 1, mode$Value, NA),
			  ifelse(mode$Freq != 1, mode$Freq, NA),
			  sd(values, na.rm = T),
			  var(values, na.rm = T),
			  quantile(values, probs = c(0.25), na.rm = T),
              quantile(values, probs = c(0.75), na.rm = T),
              quantile(values, probs = c(0.10), na.rm = T))
	)
}

summary_cols <- c("Mean", "Median", "Range", "StdDev", "Variance", "Q1", "Q3", "P10")

```

## 1.) Cars

> __Exercise 1__: mileage.csv is derived from a 1991 U.S EPA study of passenger car mileage. This file includes
information on sixty cars: HP (engine horsepower), MPG (average miles per gallon) WT (vehicle weight in
100 lb units) and CLASS (vehicle weight class C1,.,C6).

+ __Dataset__: _milage.csv_

```{r ex_3_1.data, echo = FALSE}

milage <- loadDataFile("mileage.csv")

milage$Class <- factor(milage$Class, ordered = TRUE)

glimpse(milage)

```

### 1.1) Millage Summary

> For each weight class determine the mean and standard deviation of MPG. What can you conclude from these calculations?

```{r ex_3_1.mpg_summary, echo = FALSE}

mpg_by_class <- milage[, .(Avg = mean(Mpg), StdDev = sd(Mpg)), by = Class]

print(mpg_by_class)

```


```{r ex_3_1.gg, echo = FALSE}

ggplot(milage, aes(x = Class, y = Mpg)) +
	geom_boxplot() +
	labs(title = "MPG By Class")

```

> __Answer__: There is a strong inverse correlation to MPG by weight class. This is an intuitive result, as one would expect the more a car weights, the less fuel efficient it is.
There are minimal outliers in this data set, so we should have high relatively confidence in these results.\

### 1.2) Horsepower Summary
> For each weight class determine the mean and standard deviation of HP. What can you conclude from these calculations?

```{r ex_3_2.hp_summary, echo = FALSE}

hp_by_class <- milage[, .(Avg = mean(Hp), StdDev = sd(Hp)), by = Class]

print(hp_by_class)

```


```{r ex_3_2.gg, echo = FALSE}

ggplot(milage, aes(x = Class, y = Hp)) +
	geom_boxplot() +
	labs(title = "Hoursepower By Class")

```

> __Answer__: There is a strong correlation from Horsepower to Weight Class. This makes intuitive sense as one would expect higher displacement engines, 
which typically have more horsepower, are housed in larger vehicles. This would of course not always be true, as race cars and specialty vehicles would be outliers
in this data set. We also see many outliers in the C5 class, which suggest this could fit the profile of such a race car or special purpose vehicle.

## 2.) Shoppers

> __Exercise 2:__: shoppers.csv contains the dollar amounts spent in a store by individual shoppers during one
day.

```{r ex_3_2_data, echo = FALSE}

shoppers <- loadDataFile("shoppers.csv")

glimpse(shoppers)

```

> Find the mean, median, range, standard deviation, variance, Q1, Q3 and P10. Plot the histogram and
describe the distribution.


```{r ex_3_2_summary, echo = FALSE}

shoppers.summary <- getSummary(shoppers$Spending)

print(shoppers.summary[which(Statistic %in% summary_cols)])

```


```{r ex_3_2_vis, echo = FALSE}

ggplot(shoppers, aes(Spending, fill = ..count..)) +
	geom_histogram(breaks = pretty(shoppers$Spending)) +
	labs(title = "Shopping Distribution")

```

> __Answer__: The shopping distribution is heavily right-skewed, showing that most transactions were on the lower end,
with a few big spenders sloping downward pretty linearly after around ~$20 (Assuming USD).

## 3.) POTUS

> __Exercise 3:__ pontus.csv lists the ages of USA Presidents at the time of their inauguration. Also listed are
the heights of the Presidents and their opponents.

### 3.1) Summary

> Find the mean, median, range, standard deviation, Q1, Q3 and P10 of the Presidents’ ages.

```{r ex_3_3_data, echo = FALSE}

presidents <- loadDataFile("pontus.csv")

presidents.age.summary <- getSummary(presidents$Age)

print(presidents.age.summary[which(Statistic %in% summary_cols)])

```

### 3.2) Opponents

> Find the mean, median, range, standard deviation, Q1, Q3 and P10 of the heights of the Presidents
and also their opponents.


```{r ex_3_3_summary, echo = FALSE}

lapply(apply(presidents[, .(Ht, Htopp)], 2, FUN = getSummary),
    function(x) x[which(Statistic %in% summary_cols)])

```

### 3.3) Height Comparison

> Calculate the difference between each President’s height and that of his opponent. Plot a histogram
of these differences. Construct a boxplot. What do you conclude from your calculations? Why is the
difference of average heights calculated in (2) different from the average of the pairwise differences
calculated in (3)?

```{r ex_3_3_vs_hist, echo = FALSE}

presidents[, Diff := Ht - Htopp]

presidents.complete = presidents[complete.cases(presidents)]

ggplot(presidents.complete, aes(Diff, fill = ..count..)) +
    geom_histogram( breaks = pretty(presidents$Diff))

```

```{r ex_3_3_vs_box, echo = FALSE}

ggplot(presidents.complete) +
	geom_boxplot(aes("President", Ht)) +
	geom_boxplot(aes("Opponent", Htopp))

```


> __Answer__: If you only include the complete cases (excluding any NAs), then 2 and 3 are similar in that the mean difference is 0, and the majory of the
differences are between 180 - 170 [0-10].

## 4.) Old Faithful

> __Exercise 4:__ geyser.csv contains the intervals (in minutes) between eruptions of Old Faithful Geyser in
Yellowstone National Park. The data were taken on two consecutive weeks: WEEK1 and WEEK2.

> Compare the two sets of data using summary(), hist() and boxplot(). 


```{r ex_3_4_data, echo = FALSE}

geyser <- loadDataFile("geyser.csv")

geyser_pivot <- melt(geyser, measure.vars = c("Week1", "Week2"))
colnames(geyser_pivot) <- c("Week", "Minutes")

```


```{r ex_3_4_summary, echo = FALSE}

summary(geyser)

```


```{r ex_3_4_hist, echo = FALSE}

ggplot(geyser_pivot, aes(Minutes, fill = ..count..)) +
	geom_histogram(breaks = pretty(geyser_pivot$Minutes)) +
	facet_wrap(~Week)

```


```{r ex_3_4_vis, echo = FALSE}

ggplot(geyser_pivot, aes(Week, Minutes)) +
	geom_boxplot()

```

> What do you conclude?

> __Answer:__ The two datasets are similar enough in terms of range and measures of center. Week 1 has some upper bound outliers and seemed to be
more 'volatile' overall, while week 2 has more central tendency.