---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Fitting a Bivariate t-Distribution}
   \rfoot{\color{headergrey}Module 3}
   \lfoot{\color{headergrey}MSDS 451}
   \fancyfoot[C]{\rmfamily\color{headergrey}Financial and Risk Analytics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
    df_print: paged
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
             axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             legend.position = "top", legend.title = element_blank())

```

```{r pander_setup, include = FALSE}

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Problem Background
######  

The following code block computes the MLE for a bivariate t-distribution fit to CRSP returns data.

```{r, mle_bivariate_t, echo = T}

data(CRSPday, package = "Ecdat")

Y = CRSPday[, c(5,7)]

loglik = function(par) {

  mu = par[1:2]
  
  A = matrix( c(par[3], par[4], 0, par[5] ), 
              nrow=2, byrow=T )

  scale_matrix = t(A) %*% A
  
  df = par[6]
  
  -sum(log(dmt(Y,mean=mu,S=scale_matrix,df=df)))
}

A = chol(cov(Y))

start = as.vector( c(apply(Y,2,mean), A[1,1], A[1,2], A[2,2], 4) )

fit_mvt = optim( start, 
                 loglik, 
                 method="L-BFGS-B", 
                 lower=c(-0.02,-0.02,-0.1,-0.1,-0.1,2), 
                 upper=c(+0.02,+0.02,+0.1,+0.1,+0.1,15), 
                 hessian=T )

params = fit_mvt$par
params.disp <- round(params, 7)

```
\newpage

Let $\hat{\theta} = (\mu_{1}, \mu_{2}, A_{1,1}, A_{1,2}, A_{2,2}, \upsilon)$, where
$\mu_j$ is the mean of the $ijth$ variable, $A_1, A_2, and A_3$ are the nonzero element of A, and $\upsilon$ is
the degrees of freedom parameter.

### Problem 1
###### _What does the code A = chol(cov(Y)) do_?

__cov(Y)__ computes the covariate matrix for _IBM_ and _CRSP_ returns in the CSPday dataset (from the Ecdat package).

The covariate matrix is then passed into __chol__, which computes the "Square Root" of the cov matrix using the Cholesky factorization method.

The resulting 2x2 matrix (since we have 2 stocks in this data set) is then stored in __A__.

### Problem 2
###### Find $\hat{\theta}_{ML}$, the MLE of $\hat{\theta}$.

\begin{table}[H]
\centering
\caption{$\theta_{ML}$ parameter estimates}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
$\mu_1$ & $\mu_2$ & $A_1$ & $A_2$ & $A_3$ & $v$ \\ \hline
$`r format(params.disp[1], scientific=F)`$ & $`r format(params.disp[2], scientific=F)`$ & $`r format(params.disp[3], scientific=F)`$ & $`r format(params.disp[4], scientific=F)`$ & $`r format(params.disp[5], scientific=F)`$ & $`r format(params.disp[6], scientific=F)`$ \\ \hline
\end{tabular}
\end{table}


### Problem 3
###### Find the MLE of the covariance matrix of the returns.


```{r, prob_3, echo = T}

A = matrix( c(params[3], params[4], 0, params[5] ), 
            nrow=2, byrow=T )

df <- params[6]
M.cov <- df / ( df - 2 ) * crossprod(A)

kable(format(round(M.cov, 7), scientific = F), "latex", caption = "Covariate Matrix", booktabs = T) %>%
  kableExtra::kable_styling(latex_options = "hold_position")

```

\newpage

### Problem 4
###### Find the MLE of p, the correlation between the two returns ($Y_1  and  Y_2$).


```{r, prob_4, echo = T}

p <- M.cov[1,2] / sqrt(M.cov[1,1]*M.cov[2,2])

M.cor <- matrix(c(1, p, p, 1), nrow = 2)

kable(format(round(M.cor, 7), scientific = F), caption = "MLE p, Correlation Matrix") %>%
  kableExtra::kable_styling(latex_options = "hold_position")

```
