---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Fitting Copula Models to Bivariate Returns}
   \rfoot{\color{headergrey}Module 4}
   \lfoot{\color{headergrey}MSDS 451}
   \fancyfoot[C]{\rmfamily\color{headergrey}Financial and Risk Analytics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document:
    df_print: paged
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
             axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             legend.position = "top", legend.title = element_blank())

data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Problem Background
######  Fitting Copula Models

In this lab we are going to fit copula models to a bivariate data set of daily returns on IBM and S&P500 Index.

First, we need to fit a model with the univariate marginal t-distributions and a t-copula.

This model will have three degrees of freedom parameters:

+ IBM tail index
+ S&P 500 Index tail index
+ Joint tail index (the copula)

```{r, mle_bivariate_t, echo = T}

net.returns <- read.csv(paste0(data.dir, "IBM_SP500_04_14_daily_netRtns.csv"),
                        header = T)

ibm <- net.returns[, "IBM"]
sp500 <- net.returns[, "SP500"]

suppressWarnings({
  est.ibm = as.numeric( fitdistr( ibm, "t" )$estimate )
  est.sp500 = as.numeric( fitdistr( sp500, "t" )$estimate )
})

est.ibm[2] = est.ibm[2] * sqrt( est.ibm[3] / (est.ibm[3]-2) )
est.sp500[2] = est.sp500[2] * sqrt( est.sp500[3] / (est.sp500[3]-2) )

```

The univariate estimates will be used as starting values when we estimate the _meta-t_ distribution is fit by maximum liklihood. Before we do that, we need to compute an estimate of the correlation coefficent in the _t_-copula.

### Problem 1
###### _Using Kendall's tau, compute __omega__, which is the estimate of the Pearson correlation from Kendall's tau._

From __8.27__ we have Kendall's tau, $\rho_\tau$, = 

$\rho_\tau(Y_i, Y_j) = \frac{2}{\pi}arcsin(\Omega_{i,j})$.

Inverting, we derive that:

$\Omega_{i,j} = sin[\frac{\pi}{2}\rho_\tau(Y_i, Y_j)]$

```{r, pearson_cor, echo = T}

cor_tau = cor(ibm, sp500, method = "kendall")
omega = sin((pi/2) * cor_tau)
  
```

$\Omega = `r format(round(omega, 6), scientific=F)`$

\newpage

The _t_-copula using omega as the correlation parameter and 4 as the degrees of freedom:

```{r, t_copula, echo = T}

cop_t_dim2 <- tCopula(omega, dim = 2, dispstr = "un", df = 4)

```

```{r, t_copula_disp, echo = F}
cop_t_dim2
```

Now fit copulas to the uniformed-transformed data:

```{r, t_copula_fit, echo = T}

n = nrow(net.returns)

data1 = cbind( pstd( ibm, mean=est.ibm[1], sd=est.ibm[2], nu=est.ibm[3] ),
               pstd( sp500, mean=est.sp500[1], sd=est.sp500[2], nu=est.sp500[3] ) )

data2 = cbind( rank(ibm)/(n+1), rank(sp500)/(n+1) )

ft1 = fitCopula(cop_t_dim2, data=data1, method="ml", start = c(omega,4))
ft2 = fitCopula(cop_t_dim2, data=data2, method="ml", start = c(omega,4))

```

### Problem 2
###### Explain the difference between methods used to obtain the two estimates _ft1_ and _ft2_.

Here we are fitting the copulas to the uniform-transformed data. One method is parameteric (the t-distribution) and one is non-parametric (uniform).

__data1__ is computed using the probability density function for the standard t-distrubution, since we previously estimated the paramters to fit a t-distribution to the respective returns using the __fitdistr__ function above.

__data2__ is computed using the percentile rank method.

Both of these methods are calculating the cumulative density of the marginal values in IBM and SP500, respectively.

Below we will have a quick visual inspection of the densities of the two methods.

```{r, marginal_density_1, fig.height=4, echo = F}

dt1 <- data.table(IBM = data1[, 1], SP500 = data1[, 2])

p1 <- ggplot(dt1) +
  geom_histogram(aes(IBM, y = ..density..), bins = 40, alpha = .5) +
  geom_density(aes(IBM), lwd = 1)

p2 <- ggplot(dt1) +
  geom_histogram(aes(SP500, y = ..density..), bins = 40, alpha = .5) +
  geom_density(aes(SP500), lwd = 1)

grid.arrange(p1, p2, nrow = 2, top  = "data1: Marginal Densities")

```

```{r, marginal_density_2, fig.height=4, echo = F}

dt2 <- data.table(IBM = data2[, 1], SP500 = data2[, 2])

p3 <- ggplot(dt2) +
  geom_histogram(aes(IBM, y = ..density..), bins = 40, alpha = .5) +
  geom_density(aes(IBM), lwd = 1)

p4 <- ggplot(dt2) +
  geom_histogram(aes(SP500, y = ..density..), bins = 40, alpha = .5) +
  geom_density(aes(SP500), lwd = 1)

grid.arrange(p3, p4, nrow = 2, top  = "data2: Marginal Densities")

```

\newpage

###### Do the two estimates seem significantly different (_in a practical sense_)?.

```{r, t_copula_fit_disp, echo = T}

summary(ft1)

summary(ft2)

```

These two estimates are fairly close to each other for practical purposes.

However, I think method one would be a more robust estimate due to the marginals coming from a fitted t-distribution, versus a _truly uniform_ ranking.

\newpage

### Problem 3

Next, we will define a meta-_t_-distribution by specifying its _t_-copula and its univariate marginal distributions.

```{r, fit_meta-t_copula, echo = T}

mvdc_t_t = mvdc( cop_t_dim2, c("std","std"), list(
           list(mean=est.ibm[1],sd=est.ibm[2],nu=est.ibm[3]),
           list(mean=est.sp500[1],sd=est.sp500[2],nu=est.sp500[3])))

```

```{r, fit_meta-t_copula_disp, echo = T}

mvdc_t_t

```

Now we fit the meta _t_-distribution.

```{r, fit_meta_t_mle, echo = T}

start = c(est.ibm, est.sp500, ft1@estimate)
objFn = function(param) -loglikMvdc( param, cbind(ibm,sp500), mvdc_t_t)
tic = proc.time()
ft = optim(start, objFn, method="L-BFGS-B",
           lower = c(-.1,0.001,2.2, -0.1,0.001,2.2,  0.2,2.5),
           upper = c( .1,   10, 15,  0.1,   10, 15,  0.9, 15) )
toc = proc.time()
total_time = toc - tic ; total_time[3]/60

```

```{r, param_est, echo = F}

params.disp <- round(ft$par, 6)

rho <- ft$par[7]
nu <- ft$par[8]

```

### What are the estimates of the copula parameters?

For $C_{t(\Omega, v)}$

+ $\Omega = `r format(params.disp[7], scientific = F)`$
+ $\nu = `r format(params.disp[8], scientific = F)`$

### What are the estimates of the parameters in the univariate marginal distributions?

\begin{table}[H]
\centering
\caption{Marginal t-distribution estimates}
\begin{tabular}{|l|l|l|l|}
\hline
Symbol & $\hat{\mu}$ & $\hat{\sigma}$ & $\hat{\nu}$ \\ \hline
IBM & $`r format(params.disp[1], scientific=F)`$ & $`r format(params.disp[2], scientific=F)`$ & $`r format(params.disp[3], scientific=F)`$ \\ \hline
SP500 & $`r format(params.disp[4], scientific=F)`$ & $`r format(params.disp[5], scientific=F)`$ & $`r format(params.disp[6], scientific=F)`$ \\ \hline
\end{tabular}
\end{table}

### Was the estimation method maximum likelihood, semiparametric pseudo-maximum likelihood, or parametric pseudo-maximum likelihood?

Here, since we are estimating the parameters of the marginal distributions with the parameters of the t-copula at the same time, this is a maximum likelihood estimate.

### Estimate the coefficent of lower tail dependence for this copula.

From __8.21__ we have:

$\lambda_{\ell} = 2F_{t, \nu+1}\bigg\{-\sqrt{\frac{(\nu+1)(1-\rho)}{1 + \rho}}\bigg\}$

Where,

+ $\nu = `r format(params.disp[7], scientific = F)`$
+ $\rho = `r format(params.disp[8], scientific = F)`$

```{r, lower_tail_dependence, echo = T}

threshold <- .01

# Method 1: book formula

x <- -sqrt( ( nu+1 ) * ( 1-rho ) / ( 1+rho ) )
lambda <- 2 * pt( x, nu + 1)

# Method 2: package function
lambda2 <- fitLambda(data1, method = c("t"), p = min( 100 / nrow(data1), 0.1),
          lower.tail = TRUE, verbose = F)

# assert similarity
stopifnot((lambda - lambda2$Lambda[1,2]) < threshold)

```

$\lambda_{\ell} = `r format(round(lambda, 6), scientific = F)`$