---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Efficient Equity Portfolios}
   \rfoot{\color{headergrey}Module 8}
   \lfoot{\color{headergrey}MSDS 451}
   \fancyfoot[C]{\rmfamily\color{headergrey}Financial and Risk Analytics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(quadprog, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Problem Background

Write an R program to find the efficient frontier, the tangency portfolio, and the minimum variance portfolio, and plot on "risk-reward space" the location of each of the six stocks, the efficent frontier, the tangency portfolio, and the line of efficient portfolios.

Use the constraints that, $-0.1 \le w_j \le 0.5$ for each stock.

The first constraint limits the short sales but does not rule them out completely.

The second constraint prohibits more than 50% of the investment in any single stock.

Assume that the annual risk-free rate is 3%.

```{r, stock_prices, echo = T}

data <- data.table(read.csv(paste0(data.dir, "Stock_Bond.csv"),
                 header = T))

prices <- cbind(GM = data$GM_AC, F = data$F_AC, CAT = data$CAT_AC, UTX = data$UTX_AC,
                MRK = data$MRK_AC, IBM = data$IBM_AC)

n <- nrow(prices)

returns <- data.table(100 * (prices[2:n,] / prices[1:(n-1),] - 1))

```

```{r, pairplot, echo = F, fig.height=4.5, fig.width=8}
ggpairs(returns)
```

```{r echo = T}

cov.mat <- cov(returns)
mean.vec <- colMeans(returns)
sd.vec <- sqrt(diag(cov.mat))

rfr <- 3.0 / 365 # Daily risk-free rate

n.sims <- 500 # Simulations to find optimal allocation.
n.stocks <- ncol(prices) # Stocks to allocate.
c.vec <- c(-.10, .50) # Allocations between -10% and 50%.

# Storage.

mu_p <- seq(min(mean.vec), max(mean.vec), length = n.sims)
sd_p <- mu_p

out.weights <- matrix(0, nrow = n.sims, ncol = n.stocks)

# ?solve.QP

A.mat <- cbind(rep(1, n.stocks), mean.vec)
b.vec <- c(1, NaN)

# Lower-bound
A.mat <- cbind(A.mat, diag(1, n.stocks))
b.vec <- c(b.vec, c.vec[1]*rep(1, n.stocks))

# Upper-bound
A.mat <- cbind(A.mat, -diag(1, n.stocks))
b.vec <- c(b.vec, -c.vec[2]*rep(1, n.stocks))

# Find the optimal portfolios for each target expected return.
for ( i in 1:n.sims )
{
   # constraint vector
   b.vec[2] = +mu_p[i]
  
   result =
      solve.QP( Dmat = 2*cov.mat, 
                dvec = rep(0, n.stocks),
                Amat = A.mat, 
                bvec = b.vec, 
                meq = 2)
  
   sd_p[i] = sqrt(result$value)
   out.weights[i, ] = result$solution
}

# Find maximum Sharpe's ratio
sharpe.ratios <- ( mu_p - rfr) / sd_p

# Find & save the tangency portfolio
tangent.index <- which.max(sharpe.ratios)

dt.tan.weights <- data.table(matrix(out.weights[tangent.index,], nrow = 1))
colnames(dt.tan.weights) <- colnames(prices)
      
# Find & save the minimum variance portfolio
minvar.index <- which.min(sd_p)

dt.minvar.weights <- data.table(matrix(out.weights[minvar.index,], nrow = 1))
colnames(dt.minvar.weights) <- colnames(prices)

# Portfolio weights must sum to 100%.
stopifnot(round(sum(dt.tan.weights), 4) == 1 &
          round(sum(dt.minvar.weights), 4) == 1)

# Find Sharpe Ratio of tangent portfolio.
tangent.sharpe <- sharpe.ratios[tangent.index]

```

```{r echo=T, fig.height=8, fig.width=8}

# Plot Efficent Frontier

ef.data <- data.table(cbind(sd_p, mu_p))[, Index := .I]
ef.data[, ef := ifelse(Index > minvar.index, mu_p, NA)]

stocks.data <- data.table(x = sd.vec, y = mean.vec, symbol = colnames(prices))

suppressWarnings(print({
   ggplot(ef.data) +
      
      geom_abline(aes(intercept = rfr, slope = tangent.sharpe), col = "cornflowerblue", lty = 4, lwd = 1) +
      
      geom_point(aes(sd_p, mu_p), col = "darkred", lwd = .15, alpha = .7) +
      geom_point(aes(sd_p, ef), col = "darkgreen", lwd = .15, alpha = .7) +
      
      geom_point(aes(0, rfr), col = "darkgreen", size = 3) +
      geom_text_repel(data = ef.data[tangent.index],
                  aes(0, rfr, label = "Risk-free Rate"),
                  size = 4, box.padding = 1.5,
                  force = 15, segment.size  = 0.2,
                  segment.color = "grey50", direction = "x") +
   
      geom_point(data = ef.data[tangent.index], aes(sd_p, mu_p), col = "lightblue", size = 3) +
      geom_text_repel(data = ef.data[tangent.index],
                  aes(sd_p, mu_p, label = "Tangency"),
                  size = 4, box.padding = 2,
                  force = 30, segment.size  = 0.2,
                  segment.color = "grey50",
                  direction     = "x") +
   
      geom_point(data = ef.data[minvar.index], aes(sd_p, mu_p), col = "darkgrey", size = 3) +
      geom_text_repel(data = ef.data[minvar.index],
                  aes(sd_p, mu_p, label = "Minimum Variance"),
                  size = 4, box.padding = 1.5, force = 15,
                  segment.size  = 0.2, segment.color = "grey50",
                  direction = "x") +

      geom_text_repel(data = ef.data[tangent.index - floor((tangent.index-minvar.index)/2),],
                  aes(sd_p, mu_p, label = "Efficent Frontier"),
                  size = 4, box.padding = 1.5,
                  force = 20, segment.size  = 0.2,
                  segment.color = "grey50",
                  direction = "x") +
   
      geom_text(data = stocks.data, aes(x, y, label = symbol, col = symbol)) +
      
      geom_text(data = stocks.data, aes(mean(x), mean(y), nudge_y = -.5,
                                        label = "Individual Stocks"), color = "black", size = 5, alpha = .2) +
      
      scale_y_continuous(limits = c(0, .1), labels = scales::percent_format(scale = 100)) +
      scale_x_continuous(limits = c(0, 2.5)) +
      
      labs(x = "Risk", y = "Return") +
      guides(color = "none") +
      theme(axis.line = element_line(colour = "black"),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         panel.border = element_blank(),
         panel.background = element_blank())
}))

```

```{r echo=F}

pretty_kable(cbind(Portfolio = c("Tangency", "Minimum Variance"), rbind(dt.tan.weights, dt.minvar.weights)), 
             "Optimal Portfolio Weights", dig = 5)
```