---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Fama-French Model}
   \rfoot{\color{headergrey}Module 9}
   \lfoot{\color{headergrey}MSDS 451}
   \fancyfoot[C]{\rmfamily\color{headergrey}Financial and Risk Analytics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(quadprog, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Problem Background

In this lab we are going to start with a one-factor CAPM model, and then extend it to the three-factor Fama-French model.

We will use the data set for Stocks, FX and Bonds from 2004-2005.

Data for the Fama-French factors are avilable Prof. Kenneth French's website.

Where RF is the risk-free rate and Mkt.RF, SMB and HML are the Fama-French factors.

```{r, factor_model, echo = T}

# Stock/Bond/FX data.
stocks <- as.data.table(read.csv(paste0(data.dir, "Stock_FX_Bond_2004_to_2005.csv"), 
                                 header=T))
stocks$Date <- as.Date(stocks$Date, format = "%d-%b-%y")
stocks_subset <- stocks[, .(Date, GM_AC, F_AC, UTX_AC, MRK_AC)]
stocks_diff <- data.table(Date = stocks_subset[-1]$Date,
                          100 * apply(log(stocks_subset[, .(GM_AC, F_AC, UTX_AC, MRK_AC)]), 2, diff))

# Fama-French data.
FF_data <- as.data.table(read.table(paste0(data.dir, "FamaFrenchDaily.txt"), 
                                    header=T))
FF_data$Date <- as.Date(as.character(FF_data$date), format = "%Y%m%d")
FF_data$date <- NULL

# Combine into one data.table.
consolidated.data <- merge(stocks_diff, FF_data, on = c("Date"))
capm.data <- consolidated.data[, .(Date,
                              GM = GM_AC - RF,
                              Ford = F_AC - RF,
                              UTX = UTX_AC - RF,
                              Merck = MRK_AC - RF,
                              Mkt.RF, SMB, HML, RF)]


fit <- lm(as.matrix(cbind(GM, Ford, UTX, Merck)) 
                   ~ Mkt.RF + SMB + HML, data = capm.data)
```

\newpage

### 1.)  Which one(s) of the 4 stocks are "value" stocks, according to this model? How can you tell?

```{r, echo=F}

dt.ff <- as.data.table(coef(fit), keep.rownames = T)
colnames(dt.ff)[1] <- 'Factor'
dt.ff[1, 1] <- 'Alpha'
pretty_kable(dt.ff, "Fama-French Model", dig = 5)
```

A "value" stock is one with a positive HML (high minus low) coefficent, which would be __GM__ and __Ford__. __Merck__ looks like a _"growth"_ stock, and UTX is somewhere in between a "growth" and "value" stock.

### 2.) Fama-French model fits which stock best? Worst?

```{r, echo = T}
lm.stats <- summary(fit)

dt.stats <- data.table(stat = c("R-Squared", "Adj. R-Squared"), 
            GM = c(lm.stats$`Response GM`$r.squared, 
                   lm.stats$`Response GM`$adj.r.squared),
            Ford = c(lm.stats$`Response Ford`$r.squared,
                     lm.stats$`Response Ford`$adj.r.squared),
            UTX = c(lm.stats$`Response UTX`$r.squared,
                    lm.stats$`Response UTX`$adj.r.squared),
            Merck = c(lm.stats$`Response Merck`$r.squared,
                      lm.stats$`Response Merck`$adj.r.squared))
dt.stats[, 2:5] <- dt.stats[, 2:5] * 100
pretty_kable(dt.stats, "Model Fit Statistics", dig = 2)

```

The Fama-French model fits __UTX__ the best, explaining roughtly 35% of the variance of the returns, and __Merck__ the worst, explaining about 6.6% of the return variance.

### 3.) For UTX, which factor is most important besides the market return?

For __UTX__ the most important factor outside the market return is SMB, or small minus big, at _-30%_. A strong negative relationship with SMB means it behaves, unsurprisingly, like a large cap stock.

\newpage

### 4.) Use the Fama-French model to predict the next day return of UTX. 

Use the first 250 days (__the training set__) to fit the model. Note that this model will be different from the one above. Use this model to make 1-day forecasts for the rest of the data (__the test set__). 

```{r, model_train, echo=T}

utx.data <- capm.data[, .(Date, UTX, Mkt.RF, SMB, HML)]
train.size <- 250
test.size <- nrow(utx.data) - train.size

train.data <- utx.data[1:train.size,]
test.data <- utx.data[train.size+1:test.size,]

stopifnot(nrow(train.data) == train.size & nrow(test.data) == test.size)

summary(utx.fit <- lm(UTX ~ Mkt.RF + SMB + HML,
                        data = train.data))

```

\newpage

+ What is the percentage of days in the test set where the sign of the forecast return agrees with that of the actual return? (sign=-1,0,1)

```{r, echo=T}
# Use the 1 day lagged prediction.
# i.e, the sign of 1/3/05 close is using the FF data from 12/31/04.
results <- data.table(Actual = test.data[-1]$UTX, 
                      Pred = predict(utx.fit, newdata = head(test.data, -1)))
results[, correct := sign(Actual) == sign(Pred)]
results$Date <- test.data[-1]$Date

pct.correct <- (nrow(results[correct == T]) / nrow(results)) * 100

```

Using the 3-factor model, we can correctly predict the sign of the next day return: __`r round(pct.correct, 2)`%__

```{r, model_forecast, echo = T, fig.width=8}

suppressWarnings(print({
   ggplot(results, aes(x = Date)) +
      geom_line(aes(y = Actual)) +
      geom_line(aes(y = ifelse(correct == T, Actual, NA)), col = "darkgreen") +
      geom_line(aes(y = ifelse(correct == F, Actual, NA)), col = "darkred") +
      labs(title = "Fama-French 3-Factor Prediction")
}))

```