---
title: "Lesson 09: Statistical Inference Hypothesis Testing for Single Populations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
  base.dir <- lp.w
} else if (file.exists(lp.h)) {
  base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
  data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

  data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

  return(data)
}

```

## Analysis of Variance

```{r data_tableware}
tableware <- loadDataFile("tableware.csv")

str(tableware)

```

### 1.) Rate ~ Type

Use the tableware.csv data to test the hypothesis that the mean RATE for the five levels of TYPE are
equal. Test the hypothesis using a 0.05 significance level. Plot means and confidence intervals of RATE
for each level of TYPE (Use the example given in Davies Chapter 9.3 One-way ANOVA (pp. 218-223)).

```{r}

aov_type = aov(Rate ~ Type, data = tableware)
summary(aov_type)

lm_type <- lm(Rate ~ Type, data = tableware)
summary(lm_type)

plotmeans(Rate ~ Type, data = tableware)

byType <- tableware[, .(Rate.mean = mean(Rate), Rate.sd = sd(Rate), N = length(Rate), tfrac = qt(p = .975, df = length(Rate) - 1)), by = list(Type)][,
	":="(Lower = Rate.mean - tfrac * Rate.sd / sqrt(N), Upper = Rate.mean + tfrac * Rate.sd / sqrt(N))]

ggplot(byType, aes(x = Rate.mean, y = Type)) +
	geom_point(color = "steelblue", size = 3) +
	geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = .3) +
    ggtitle("Average Rate by Type")

```

```{r}

```

### 2.) Price ~ Type

Use the tableware.csv data to test the hypothesis that the mean PRICE for the five levels of TYPE are
equal. Test the hypothesis using a 0.05 significance level. Plot means and confidence intervals of PRICE
for each level of TYPE (Use the example given in Davies Chapter 9.3 One-way ANOVA (pp. 218-223)).

```{r}

price_model <- { Price ~ Type }

price_fit <- lm(price_model, data = tableware)

print(summary(price_fit))

print(anova(price_fit))

print(confint(price_fit, level = 0.95))

```

### 3.)

Use the hot_dogs.csv data. Perform a one-way AOV by Type on Calories and also Sodium (Use the
example given in Davies Chapter 9.3 One-way ANOVA (pp. 218-223)). Use Tukey’s Honest Significant
Difference Test if the F-test is significant. Generate boxplots.

```{r}
hotdogs <- loadDataFile("hot_dogs.csv")

ggplot(hotdogs, aes(x = Type, y = Calories, group = Type)) +
	geom_boxplot(fill = "steelblue")

ggplot(hotdogs, aes(x = Type, y = Sodium, group = Type)) +
	geom_boxplot(fill = "steelblue")

calories.anova <- aov(Calories ~ Type, data = hotdogs)
summary(calories.anova)

sodium.anova <- aov(Sodium ~ Type, data = hotdogs)
summary(sodium.anova)

# Perform Tukey's Honest Significant Difference Test
TukeyHSD(calories.anova, conf.level = 0.95)

```
