---
title: "Lesson 02: Charts and Graphs"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
	base.dir <- lp.w
} else if (file.exists(lp.h)) {
	base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

```

> __Data Set:__

```{r data-prep, echo = FALSE}

data.path <- "C:/Projects/R/Playground/Classes/MSDS401/Exercises/02_charts/data/"

data.raw <- as.data.table(fread(paste0(data.path, "home_prices.csv"),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", "")))

home.sales <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

home.sales$Year <- as.Date(paste(home.sales$Year, 12, 31, sep = "-"))
home.sales$Baths <- factor(home.sales$Baths, ordered = TRUE)
home.sales$Feats <- factor(home.sales$Feats, ordered = TRUE)
home.sales$Nbr <- home.sales$Nbr == "YES"
home.sales$Corner <- home.sales$Corner == "YES"

glimpse(home.sales)

```

### Exercises 2.1

> __Exercises: 1__: For the following exercises use hist(), plot(), boxplot() and par() functions supplied by R.

##### a.) Construct a histogram for PRICE. Describe the distribution shape.

+ _Base R:_

```{r ex_2.1_a_br, echo = FALSE}

br.price_hist <- function() {
    with(home.sales, hist(Price,
        breaks = pretty(home.sales$Price),
        main = "Distribution of Home Prices",
        xlab = "Price", ylab = "Frequency",
        col="blue"))
}

br.price_hist()

```

+ _ggplot2:_

```{r ex_2.1_a_gg, echo = FALSE}

gg.price_hist <- ggplot(home.sales, aes(Price, fill = ..count..)) +
	geom_histogram(breaks = pretty(home.sales$Price)) +
	scale_x_continuous(labels = scales::dollar) +
	labs(title = "Distribution of Home Prices") +
	theme(plot.title = element_text(hjust = 0.5)) +
	theme(axis.text.x = element_text(face = "bold", size = 12)) +
	theme(axis.text.y = element_text(face = "bold", size = 12)) +
	theme(legend.position = "top", legend.title = element_blank())

gg.price_hist

```

> __Answer__: The distribution has a right skewed distibution, indicating that more homes were sold with lower prices.

##### b) Construct a histogram for TAX. Describe the distribution shape.

+ _Base R:_

```{r ex_2.1_b_br, echo = FALSE}

br.tax_hist <- function() {
    with(home.sales,hist(Tax,
        breaks = pretty(home.sales$Tax),
        main = "Distribution of Home Taxes",
        xlab = "Tax", ylab = "Frequency",
        col="orange"))
}

br.tax_hist()

```

+ _ggplot2:_

```{r ex_2.1_b_gg, echo = FALSE}

gg.tax_hist <- ggplot(home.sales, aes(Tax, fill = ..count..)) +
	geom_histogram(breaks = pretty(home.sales$Tax)) +
	scale_x_continuous(labels = scales::dollar) +
	theme(plot.title = element_text(hjust = 0.5)) +
	theme(axis.text.x = element_text(face = "bold", size = 12)) +
	theme(axis.text.y = element_text(face = "bold", size = 12)) +
	theme(legend.position = "top", legend.title = element_blank()) +
    labs(title = "Distribution of Home Taxes")

gg.tax_hist

```

##### c) Construct a scatterplot displaying TAX versus PRICE. Is there a relationship?

+ _Base R:_

```{r ex_2.1_c_br, echo = FALSE}

plot(home.sales$Price, home.sales$Tax,
  main = "Tax vs. Price",
  xlab = "Tax", ylab = "Price",
  bg = "black", pch = 16)

# Offset Y to Origin, rough manual adjustment
abline(lm( I( Tax ) ~ I( Price ), data = home.sales), col = "red")

```

+ _ggplot2:_

```{r ex_2.1_c_gg, echo = FALSE}

ggplot(home.sales, aes(Tax, Price)) +
	geom_point() +
	geom_smooth( method = "lm") +
	scale_x_continuous(labels = scales::dollar) +
	scale_y_continuous(labels = scales::dollar) +
	labs(title = "Tax vs. Price") +
	theme(plot.title = element_text(hjust = 0.5)) +
	theme(axis.text.x = element_text(face = "bold", size = 12)) +
	theme(axis.text.y = element_text(face = "bold", size = 12)) +
	theme(legend.position = "top", legend.title = element_blank())

```

> __Answer:__ The relationship between TAX and PRICE is approximately linear.

##### d) Construct a stem-and-leaf plot for TAX using stem().

+ _Base R:_

```{r ex_2.1_d_br, echo = FALSE}

stem(home.sales$Tax)

```

+ _ggplot2:_

```{r ex_2.1_d_gg, echo = FALSE}

ggplot(home.sales, aes(x = -Tax)) +
	geom_histogram(bins = 21) +
	coord_flip() +
	labs(title = "Graphical Distribution") +
	theme(plot.title = element_text(hjust = 0.5)) +
	theme(axis.text.x = element_blank()) +
    theme(axis.text.y = element_blank())

```

##### e) Use the par() and mfrow() or mfcol() functions to construct a multi-plot figure of two rows and one column showing the histograms for PRICE and TAX.

+ _Base R:_

```{r ex_2.1_e_br, echo = FALSE}

opar = par(mfrow = c(2, 1), mar = c(2, 1, 1, .5))

br.price_hist()
br.tax_hist()

par(opar)

```

+ _ggplot2:_

```{r ex_2.1_e_gg, echo = FALSE}

grid.arrange(gg.price_hist, gg.tax_hist, nrow = 2)

```

### Exercises 2.2

> __Exercises: 2__: For the following exercises use hist() and, within hist(), breaks.

##### a) Construct a histogram for PRICE starting the first class at 1300 with a class width of 600.

+ _Base R:_

```{r ex_2.2_a_br, echo = FALSE}

inc <- 600
price_breaks <- seq( from = 1300, to = max(home.sales$Price) + inc, by = inc )

with(home.sales, hist(Price,
	breaks = price_breaks,
	main = "Distribution of Home Prices",
	xlab = "Price", ylab = "Frequency",
	col = "blue"))

```

+ _ggplot2:_

```{r ex_2.2_a_gg, echo = FALSE}

ggplot(home.sales, aes(Price, fill = ..count..)) +
	geom_histogram(breaks = price_breaks) +
	scale_x_continuous(labels = scales::dollar) +
	labs(title = "Distribution of Home Prices") +
	theme(plot.title = element_text(hjust = 0.5)) +
	theme(axis.text.x = element_text(face = "bold", size = 12)) +
	theme(axis.text.y = element_text(face = "bold", size = 12)) +
	theme(legend.position = "top", legend.title = element_blank())

```

##### b) Construct a histogram for TAX starting the first class at 500 with a class width of 500.

+ _Base R:_

```{r ex_2.2_b_br, echo = FALSE}

inc <- 500
tax_breaks <- seq(from = 500, to = max(home.sales$Tax) + inc, by = inc)

with(home.sales, hist(Tax,
	breaks = tax_breaks,
	main = "Distribution of Home Taxes",
	xlab = "Tax", ylab = "Frequency",
	col = "orange"))

```

+ _ggplot2:_

```{r ex_2.2_b_gg, echo = FALSE}

ggplot(home.sales, aes(Tax, fill = ..count..)) +
	geom_histogram(breaks = tax_breaks) +
	scale_x_continuous(labels = scales::dollar) +
	theme(plot.title = element_text(hjust = 0.5)) +
	theme(axis.text.x = element_text(face = "bold", size = 12)) +
	theme(axis.text.y = element_text(face = "bold", size = 12)) +
	theme(legend.position = "top", legend.title = element_blank()) +
	labs(title = "Distribution of Home Taxes")

```