---
title: 'Moretz Brandon, Programming with R Assignment #1'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r setup, include=FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)

```

```{r additional-libraries, echo=FALSE}

library(pander)

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_fivethirtyeight())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
             plot.subtitle = element_text(hjust = 0.5),
             axis.title = element_text( size = 10 ),
             axis.text.x = element_text(face = "bold", size = 12),
             axis.text.y = element_text(face = "bold", size = 12),
             legend.position = "top", legend.title = element_blank())

display_vector <- function(vec, label = "") {
  pander::pander(vec)
}

```

### Test Items

Read each question carefully and address each element.  Do not output contents of vectors or data frames unless requested.

## 1.) Vector Manipulations

(4 points) This problem deals with vector manipulations.

### a.) Concatenation

Create a vector that contains the following, in this order, and __output the final, resulting vector__.

__Do not round any values, unless requested.__

+ A sequence of integers from 0 to 4, inclusive.
+ The number 13
+ Three repetitions of the vector c(2, -5.1, -23).
+ The arithmetic sum of 7/42, 3 and 35/42

```{r test1a}

# Note: Parenthesis are not required in the final element. They are included for readability.
values <- c(seq(from = 0, to = 4), 13, rep(c(2, -5.1, -23), 3), (7 / 42) + 3 + (35 / 42))

display_vector(values)

```

### b.) Sorting

+ Sort the vector created in __(1)(a)__ in ascending order.  __Output this result__.  
+ Determine the length of the resulting vector and assign to "L". __Output L__. 
+ Generate a descending sequence starting with L and ending with 1.  
+ Add this descending sequence arithmetically the sorted vector. This is __vector addition__, _not vector combination_. __Output the contents__. 

_Do not round any values._

```{r test1b}

values.sorted <- sort(values, decreasing = F)

display_vector(values.sorted)

L <- length(values.sorted)

display_vector(L)

values.desc <- seq(from = L, to = 1)

result <- values.desc + values.sorted

display_vector(result)

```

### c.) Subsetting

+ Extract the first and last elements of the vector you have created in __(1)(b)__ to form another vector with the extracted elements. 
+ Form a third vector from the elements not extracted. 
+ __Output these vectors__.

```{r test1c}

result.fl <- c(first(result), last(result))

display_vector(result.fl)

result.exc <- result[-which(result %in% result.fl)]

display_vector(result.exc)

```

### d.) Reconstruction

+ Use the vectors from __(c)__ to reconstruct the vector in __(b)__.
+ __Output this vector__. 
+ Sum the elements and round to two decimal places.

```{r test1d}

result.d <- c(result.fl[1], result.exc, result.fl[2])

stopifnot(all.equal(result,result.d))

display_vector(result.d)

display_vector(round(sum(result.d), 2))

```

-----

## 2.) Trigonometric Function

(5 points) The expression y = sin(x/2) + cos(x/2) is a trigonometric function.

### a.) User-Defined Function

Create a user-defined function - via *function()* - that implements the trigonometric function above, accepts numeric values, "x," 
calculates and returns values "y."

```{r test2a}

trig_fun <- function(x) {
    sin(x/2) + cos(x/2)
}

```

### b.) Logical Subset

+ Create a vector, x, of 4001 equally-spaced values from -2 to 2, inclusive. 

+ Compute values for y using the vector x and your function from __(2)(a)__.  **Do not output x or y.**

+ Find the value in the vector x that corresponds to the __maximum__ value in the vector y. 
Restrict attention to only the values of x and y you have computed; i.e. do not interpolate. 
Round to 3 decimal places and output both the minimum y and corresponding x value.

+ Finding the two desired values can be accomplished in as few as two lines of code.  

__Do not use packages or programs you may find on the internet or elsewhere.__

__Do not output the other elements of the vectors x and y.__

Use coding methods shown in the *Quick Start Guide for R*.

```{r test2b}

x <- seq(from = -2, to = 2, length.out = 4001)
y <- trig_fun(x)

max_y <- max(y)
x_val <- x[y == max_y]

stopifnot( trig_fun(x_val) == max_y )

display_vector(round(c(x_val, max_y), digits = 3))

```

### c.) Visualization

+ Plot y versus x in color, with x on the horizontal axis. 
+ Show the location of the __maximum__ value of y determined in __2(b)__.  
+ Show the values of x and y corresponding to the __maximum__ value of y in the display.
+ Add a title and other features such as text annotations. 

_Text annotations may be added via *text()* for base R plots and *geom_text()* or *geom_label()* for ggplots._

```{r test2c}

trig.data <- data.table(x,y)

offset <- .125

ggplot(trig.data, aes(x, y), label = "f(x)", show.legend = T) +
	geom_line(color = "#00ba38", lwd = 2) +
	geom_point(aes(x_val, max_y), color = "#f8766d", size = 3.5) +
    annotate("text", x = x_val, y = max_y + offset, label = paste("max(Y) =", round(max_y,3 ))) + 
    labs(title = "Trigonometric Function",
       subtitle = paste("Max Y value (", round(max_y, 3), ") occures at x value (", round(x_val, 3), ")"),
       caption = "MSDS 401: R Programming 1, Q2", x = "X", y = "Y")

```

-----

## 3.) Function Intersection

_(4 points)_ This problem requires finding the point of intersection of two functions. 
Using the function `y = cos(x/2)*sin(x/2)`, 

+ find where the curved line `y = -(x/2)**3` intersects it within the range of values used in part (2) 
_(i.e. 4001 equally-spaced values from -2 to 2)_.  
+ Plot both functions on the same display, and show the point of intersection. 
+ Present the coordinates of this point as text in the display.

NOTE:  the functions to implement are enclosed in backticks. Ignore the backticks when implementing in code.


+ Plot both functions on the same display, and show the point of intersection.  
+ Present the coordinates of this point as text in the display.

```{r test3}

f1 <- function(x) {
	cos(x / 2) * sin(x / 2)
}

f2 <- function(x) {
    -(x / 2) ** 3
}

fun.values <- melt(data.table(x, y1 <- f1(x), y2 <- f2(x),
   mid <- y2 + y1)[,
        .(x, y1, y2, mid)], id = "x")


offset <- .25

fun.values$Type <- ifelse(fun.values$variable == 'mid', 'solid', 'dotdash')

fun.intercept <- fun.values[variable == "mid" & value == 0,][, .(x, y = value)]

ggplot(fun.values, aes(x = x, y = value, color = variable, linetype = Type, group = as.factor(variable))) +
	geom_line(lwd = 1.4) +
	geom_point(aes(x = fun.intercept$x, y = fun.intercept$y), size = 3, color = "#8E44AD", alpha = .8) +

	# These two lines also mark the intersection.
	geom_vline(xintercept = fun.intercept$x, linetype = 'longdash', alpha = .5, lwd = .25) +
	geom_hline(yintercept = fun.intercept$y, linetype = 'longdash', alpha = .5, lwd = .25) +

	# I think this looks better with text in the title area only, with the following line commented out.
	# but wanted to make sure I followed instructions.
	annotate("text", x = fun.intercept$x, y = fun.intercept$y + offset, parse = TRUE, label = paste0("Intersect (", fun.intercept$x, ",", fun.intercept$y, ")")) +
    scale_linetype(guide = 'none') +
    labs(title = "Function Intersection",
       subtitle = paste0("Intersection occurs at point: (", fun.intercept$x, ",", fun.intercept$y, ")"),
       caption = "MSDS 401: R Programming 1, Q3", x = "X", y = "Y")

```

-----

## 4.) Trees

(6 points) Use the "trees" dataset for the following items.

_This dataset has three variables (Girth, Height, Volume) on 31 felled black cherry trees._

### a.) Dataset

+ Use *data(trees)* to load the dataset. 
+ Check and output the structure with *str()*. 
+ Use *apply()* to return the median values for the three variables. __Output these values__.  
+ Using R and logicals, output the row number and the three measurements - Girth, Height and Volume - of any trees with Girth equal to median Girth. 

_It is possible to accomplish this last request with one line of code._ 

```{r test3a}

# same as data(trees)
trees <- as.data.table(trees)

str(trees)

# base R way
base_medians <- apply(trees, 2, median)

# data.table way
medians <- trees[, lapply(.SD, median, na.rm = TRUE)]

medians[, .(Girth, Height, Volume)] # display medians

trees[, Row := .I][Girth == medians$Girth][, .(Row, Girth, Height, Volume)] # 1 liner

```

### b.) Calculation

Girth is defined as the diameter of a tree taken at 4 feet 6 inches from the ground. 

+ Convert each diameter to a radius, r. 
+ Calculate the cross-sectional area of each tree using pi times the squared radius. 
+ Present a stem-and-leaf plot of the radii, and a histogram of the radii in color. 
+ Plot Area (y-axis) versus Radius (x-axis) in color showing the individual data points. 

_Label appropriately._

```{r test3b}

trees[, Radius := Girth / 2][, Area := pi* Radius ** 2 ]

stem(trees$Radius)

ggplot(trees, aes(Radius, fill = ..count..)) +
	geom_histogram(breaks = pretty(trees$Radius)) +
	labs(title = "Tree Radii Distribution",
	   caption = "MSDS 401: R Programming 1, Q4")

ggplot(trees, aes(x = Radius, y = Area)) +
	geom_point() +
	labs(title = "Tree Girth",
        subtitle = "Radis vs Area",
	   caption = "MSDS 401: R Programming 1, Q4")

```

### c.) Visualization

+ Present a horizontal, notched, colored boxplot of the areas calculated in (b). Title and label the axis.

```{r test3c}

ggplot(trees, aes( "", y = Area)) +
	geom_boxplot( outlier.colour = "red", fill = "#20D4EA", outlier.shape = 1, notch = T) +
    coord_flip() +
	labs(title = "Tree Area",
	   caption = "MSDS 401: R Programming 1, Q4") +
   	theme(axis.text.x = element_blank())

```

### d.) Statistical Test

+ Demonstrate that the outlier revealed in the boxplot of Volume is not an extreme outlier.  

_It is possible to do this with one line of code using *boxplot.stats()* or 'manual' calculation and logicals._

+ Identify the tree with the largest area and output on one line its row number and three measurements.

```{r test3d}

ggplot(trees, aes("", y = Volume)) +
	geom_boxplot(outlier.colour = "red", outlier.shape = 1, fill = "#20D4EA") +
	coord_flip() +
	labs(title = "Tree Volume",
	   caption = "MSDS 401: R Programming 1, Q4") +
   	theme(axis.text.x = element_blank())

# manual calculation:

ex_outlier_coef <- 3

q1 <- quantile(trees$Volume, .25)
q2 <- quantile(trees$Volume, .75)

ex_outlier_range <- q2 + ex_outlier_coef * (q2 - q1)

max_vol <- max(trees$Volume)

max_vol > ex_outlier_range # Not extreme.

# boxplot.stats method.

vol_extremes <- boxplot.stats(trees$Volume, coef = ex_outlier_coef)
vol_extremes$out # Not extreme.

head(trees[order(-trees$Area)][, .(Row, Girth, Height, Volume)], 1) # 1 liner

```

-----

## 5.) Exponential Distribution

(6 points) The exponential distribution is an example of a right-skewed distribution with outliers.  
This problem involves comparing it with a normal distribution which typically has very few outliers.  

### a.) Distribution

+ Use *set.seed(124)* and *rexp()* with n = 100, rate = 5.5 to generate a random sample designated as y.  
+ Generate a second random sample designated as x with *set.seed(127)* and *rnorm()* using n = 100, mean = 0 and sd = 0.15.  
+ Generate a new object using *cbind(x, y)*. __Do not output this object__; instead, assign it to a new name. 
Pass this object to *apply()* and compute the inter-quartile range (IQR) for each column:  x and y. 
+ Use the function *IQR()* for this purpose. __Round the results to four decimal places and present__ (this exercise shows the similarity of the IQR values.).

_For information about *rexp()*, use *help(rexp)* or *?rexp()*.  **Do not output x or y.**_

```{r test5a}

set.seed(124)
y <- rexp(n = 100, rate = 5.5)

set.seed(127)
x <- rnorm(n = 100, mean = 0, sd = 0.15)

values <- cbind(x, y)

sapply(apply(values, 2, IQR), FUN = function(val) { round( val, 4 )})

```

### b.) Visualization

This item will illustrate the difference between a right-skewed distribution and a symmetric one. 

+ For base R plots, use *par(mfrow = c(2, 2))* to generate a display with four diagrams; *grid.arrange()* for ggplots. 
+ On the first row, for the normal results, present a histogram and a horizontal boxplot for x in color.  
+ For the exponential results, present a histogram and a horizontal boxplot for y in color.

```{r test5b}

values <- as.data.table(values)

p1 <- ggplot(values, aes(x = x, fill = ..count..)) +
	geom_histogram(breaks = pretty(values$x))

p2 <- ggplot(values, aes("", x)) +
	geom_boxplot(outlier.colour = "red", outlier.shape = 1, fill = "#20D4EA") +
	coord_flip() +
	theme(axis.text.x = element_blank())

p3 <- ggplot(values, aes(x = y, fill = ..count..)) +
	geom_histogram(breaks = pretty(values$y))

p4 <- ggplot(values, aes("", y)) +
	geom_boxplot(outlier.colour = "red", outlier.shape = 1, fill = "#20D4EA") +
	coord_flip() +
	theme(axis.text.x = element_blank())

grid.arrange(p1, p2, newpage = TRUE, p3, p4, top = "Distributions")

```

### c.) QQ-Plot

QQ plots are useful for detecting the presence of heavy-tailed distributions. 

+ Present side-by-side QQ plots, one for each sample, using *qqnorm()* and *qqline()*.  
+ Add color and titles. In base R plots, "cex" can be used to control the size of the plotted data points and text.
+ Lastly, determine if there are any extreme outliers in either sample.

```{r test5c}

# base R method. This is to ensure I followed instructions.

opar = par(mfrow = c(1, 2))

qqnorm(values$x, col = "black",  main = "Normal Distribution", xlab = "", ylab = "sample")
qqline(values$x, lty = 2, lwd = 2, col = "blue")

qqnorm(values$y, col = "black", main = "Exponential Distribution", xlab = "", ylab = "sample")
qqline(values$y, lty = 2, lwd = 2, col = "blue")

mtext("QQ Plot, Normal & Exponential", side = 3, line = -1.5, outer = TRUE)

par(opar)

# ggplot method, this is my perfered method which I believe to be identical.

#' Title
#' geom_qqline
#' @param data - vector of data used to generate line.
#' @param color - color of the qq line
#' @param line_width - width of the qq line
#' @param alpha - transparency of the qq line
#' @param line_type - qq line display type
#'
#' @return geom object that can be used in ggplot layering. This should be identical to the qqline in base R,
#' based upon by understanding of ?qqline
#' @export
#'
#' @examples see below:
geom_qqline <- function(data,
    color = "red", line_width = 1, alpha = 1, line_type = "dashed") {

    y <- quantile(data[!is.na(data)], c(0.25, 0.75))
	x <- qnorm(c(0.25, 0.75))

	# calculate slope and y-intercept
	slope <- diff(y) / diff(x)
	int <- first(y) - slope * first(x)

	# qq-line, in ggplot form.
  geom_abline(slope = slope, intercept = int, color = color, lwd = line_width, alpha = alpha, linetype = line_type)
}

q1 <- ggplot(values, aes(sample = x)) +
	stat_qq() +
	geom_qqline( values$x, color = "blue", alpha = .8 ) +
	labs(title = "Normal Distribution", x = "")

q2 <- ggplot(values, aes(sample = y)) +
	stat_qq() +
	geom_qqline( values$y, color = "blue", alpha = .8 ) +
	labs(title = "Exponential Distribution", x = "")

grid.arrange(q1, q2, ncol = 2, top = "QQ Plot, Normal & Exponential")

# Detect any extreme outliers in x or y.
x_extremes <- boxplot.stats(values$x, coef = ex_outlier_coef) # no extreme outliers
x_extremes$out

y_extremes <- boxplot.stats(values$y, coef = ex_outlier_coef) # 1 extreme outlier, 
y_extremes$out

```
