---
title: "Lesson 09: Statistical Inference Hypothesis Testing for Single Populations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
       axis.text.x = element_text(size = 10),
       axis.text.y = element_text(size = 10),
       axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
       legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig, big.mark = "'") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
  pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
  base.dir <- lp.w
} else if (file.exists(lp.h)) {
  base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
  data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

  data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

  return(data)
}

```

# Questions

## 1.) Drug Effectiveness

> A double-blind clinical trial of a new drug for back pain was designed using control and treatment
groups. Volunteers were fully informed and assigned at random to each group. Neither the volunteers
nor the doctor knew when the new drug or a placebo was being administered. When 100 volunteers
in each group had been treated and evaluated, the results revealed an 85% success rate for the new
drug and a 65% success rate for the control group. At the 95% confidence level, is there a statistically
significant difference between the two reported rates? Use a one-sided test. Also, report a confidence
interval for the difference.

```{r question1}

x <- matrix(c(85, 65, 15, 35), nrow = 2, ncol = 2, byrow = FALSE,
    dimnames = list(c("new_drug", "control"), c("success", "fail")))

pretty_kable(x, "Trial Results")

test.results <- prop.test(x, alternative = "greater", conf.level = 0.95, correct = FALSE)

test.results

ifelse(test.results$p.value < 0.05, "Reject Null", "Cannot Reject Null")

```

## 2.)

> Two baseball players had their career records compared. In 267 times at bat, one player hit 85 home
runs. In 248 times at bat, the other player hit 89 home runs. Assume the number of home runs follows
a binomial distribution, is there a statistically significant difference with 95% confidence between the
home run averages for these two baseball players?

```{r question2}

```

## 3.)

> Using the home_prices.csv data (described in Lesson 1), compare mean selling prices between homes
located in the northeast sector of the city versus the remaining homes. Also, compare the mean selling
prices between homes with a corner lot and those located elsewhere. Use two-sample t-tests for the
hypothesis tests at the 95% confidence level. Report confidence intervals for each.

```{r question3}

home.prices <- loadDataFile("home_prices.csv")


```

## 4.)

> The nsalary.csv data are derived from data collected by the Department of Social Services of the State
of New Mexico. The data have been adapted for this problem. Using these data compare mean salary
levels between RURAL and non-RURAL locations. Use a two-sample t-test at the 95% confidence level.
Report your results.

```{r question4}

nsalary <- loadDataFile("nsalary.csv")

```

## 5.)

> tires.csv contains data published by R.D. Stichler, G.G. Richey, and J. Mandel, “Measurement of
Treadware of Commercial Tires, Rubber Age, 73:2 (May 1953). Treadwear measures of tires each tire
was subject to measurement by two methods, the first based on weight loss and the second based on
groove wear. Use a paired t-test at the 95% confidence level to test for a difference between the two
methods. Report your results using a confidence interval.

```{r question5}

```
