---
title: "Lesson 09: Statistical Inference Hypothesis Testing for Single Populations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
	base.dir <- lp.w
} else if (file.exists(lp.h)) {
	base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

```

## Simple Regression Analysis and Correlation

```{r data_tableware}

newspapers <- loadDataFile("newspapers.csv")

pretty_str <- data.table(Variable = names(newspapers),
	Classes = sapply(newspapers, typeof),
	Glimpse = sapply(newspapers, function(x) paste0(head(x), collapse = ", ")))

pretty_kable(pretty_str, "Newspapers Raw Data")

```

### 1.) Sunday vs Daily

Plot Sunday circulation versus Daily circulation. Does the scatter plot suggest a linear relationship
between the two variables? Calculate the Pearson Product Moment Correlation Coefficient between
Sunday and Daily circulation.

```{r}

ggplot(newspapers, aes(x = Daily, y = Sunday)) +
	geom_point() +
    geom_smooth(method = lm, se = T ) +
    labs( title = "Sunday vs. Daily")

with(newspapers, pretty_vector(c(Correlation = cor(Daily, Sunday))))

```

### 2.) Regression Line

Fit a regression line with Sunday circulation as the dependent variable. Plot the regression line with
the circulation data. (Use Lander pages 212 and 213 for reference.) Comment on the quality of the fit.
What percent of the total variation in Sunday circulation is accounted for by the regression line?


```{r}

lm.fit <- lm(Sunday ~ Daily, newspapers)

lm.sum <- summary(lm.fit)

ggplot(newspapers, aes(x = Daily, y = Sunday)) +
	geom_point() +
	stat_smooth(method = "lm")

par(mfrow = c(2, 2))
plot(lm.fit)
par(mfrow = c(1, 1))

```

### 3.) Confidence

Obtain 95% confidence intervals for the coefficients in the regression model. Use confint().


```{r}

confint(lm.fit, level = 0.95)

pretty_kable( predict(lm.fit, interval = "prediction", level = 0.95), "Prediction Intervals" )

Daily <- 500
Sunday <- NA
new.frame <- data.frame(Daily, Sunday)

pretty_vector( predict(lm.fit, newdata = new.frame, interval = "prediction", level = 0.95) )

```

### 4.)

Determine a 95% prediction interval to predict Sunday circulation for all available values of Daily
circulation. Use predict(model, interval=“prediction”, level=0.95). Then, define a new data frame
using Daily = 500 and Sunday = NA. Predict an interval for Sunday circulation.


```{r}

```

### 5.)

Use the tableware.csv data. Regress PRICE as a dependent variable against TIME. Comment on the
quality of the fit. Is a simple linear regression model adequate or is something more needed?


```{r}

```

### 6.)

Use the tableware.csv data. ANOVA can be accomplished using a regression model. Regress PRICE
against the variables BOWL, CASS, DISH and TRAY as they are presented in the data file. What do
the coefficients represent in this regression model? How is the effect of plate accounted for?


```{r}

```

### 7.)

Use the tableware.csv data. Plot PRICE versus DIAM and calculate the Pearson product moment
correlation coefficient. Include DIAM in the regression model in (6). Compare results between the two
models. DIAM is referred to as a covariate. Does its inclusion improve upon the fit of the first model
without DIAM?

```{r}

```

