---
title: "Lesson 08: Statistical Inference Estimation for Single Populations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = F}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```

```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
	base.dir <- lp.w
} else if (file.exists(lp.h)) {
	base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

```

```{r utility, include = F}

normal_z <- function(pct, mean = 0, sd = 1) {
    qnorm( 1 - (( 1 - pct ) / 2), mean, sd, lower.tail = T)
}

```

# Estimation for Single Populations

## 1.) Confidence Interval

> Assume a random sample of size 100 is drawn from a normal distribution with variance 1. The average
value of the sample is 50. Find a 95% confidence interval for the mean.

```{r ex_1, echo=T}

n <- 100 # sample size
mean <- 50 # mean of sample
sd <- sqrt(1) # standard deviation of population
margin.of.error <- normal_z(.95) * sd / sqrt(n)

conf.int <- c(mean - margin.of.error, mean + margin.of.error)

pretty_vector( conf.int )

```

## 2.)

> Assume the standard deviation for a normal distribution is equal to 100 units. Also assume we want
to estimate the unknown mean with a 95% confidence interval of total width 8 units. Calculate the
sample size required.


```{r ex_2, echo=T}

z_score <- qnorm(0.025, mean = 0, sd = 1, lower.tail = F)
sample_size <- sample_size <- ( z_score * 100 / 4 ) ** 2

pretty_vector(sample_size)

```

## 3.) 

> A random sample of 1600 registered voters are contacted and asked a variety of questions. For one
question, 60% of the voters expressed approval and 40% disapproval. Calculate a 95% confidence
interval for the proportion expressing approval.


```{r echo=T}

# Here we will use a built-in R function for testing a proportion.
# prop.test(x, n, p = NULL,
# alternative = c("two.sided", "less", "greater"),
# conf.level = 0.95, correct = TRUE
# x a vector of counts of successes, a one-dimensional table with two entries,
# or a two-dimensional table (or matrix) with 2 columns, giving the counts of
# successes and failures, respectively.
# n a vector of counts of trials; ignored if x is a matrix or a table.
# p a vector of probabilities of success. The length of p must be the same
# as the number of groups specified by x, and its elements must be greater
# than 0 and less than 1.
# alternative: a character string specifying the alternative hypothesis,
# must be one of "two.sided" (default), "greater" or "less". You can specify
# just the initial letter. Only used for testing the null that a single proportion
# equals a given value, or that two proportions are equal; ignored otherwise.
# conf.level: confidence level of the returned confidence interval.
# Must be a single number between 0 and 1. Only used when testing the
# null that a single proportion equals a given value, or that two proportions
# are equal; ignored otherwise.
# correct: a logical indicating whether Yates' continuity correction should be
# applied where possible.

prop_test_object <- prop.test(x = 1600 * 0.6, n = 1600,
    alternative = "two.sided", conf.level = 0.95)

str(prop_test_object)

# Notice that this object is a list and the confidence interval lower
# and uppler limits themselves can be extracted from the object as
pretty_vector( as.numeric(prop_test_object$conf.int))

```

## 4.)

> A random sample of consumers are presented with two beverages in random order and asked which
they prefer most. All the consumers expressed a preference. One beverage was preferred 85% of the
time. Use this number to determine how large a sample of consumers would be needed to generate a
95% confidence interval with an overall width just less than 2% (i.e. from 84% to 86%)?

```{r echo=T}

p <- 0.85
z_score <- qnorm(0.025, mean = 0, sd = 1, lower.tail = FALSE)
sample_size <- (z_score ** 2) * p * (1 - p) / (0.01) ** 2
round(sample_size)

sample_size <- (z_score ** 2) * (0.25) / (0.01) ** 2
round(sample_size)
 
```

# Data Analysis

> Results of a laboratory analysis of calories and sodium content of major hot dog brands.
Researchers for Consumer Reports analyzed three types of hot dog: beef, poultry, and meat (mostly pork
and beef, but up to 15% poultry meat). Fifty four observations are reported.


```{r data_hotdog}

hotdogs <- loadDataFile("hot_dogs.csv")

```

## 1.)

> Create boxplots and find 95% confidence intervals for the mean amount of calories in each Type of hot
dog: beef, meat and poultry. Construct 99% one-sided lower confidence intervals for the mean amount
of calories in each Type of hot dog: beef, meat and poultry.

```{r data_1, echo=T}

ggplot(hotdogs, aes(x = Type, y = Calories, group = Type)) +
	geom_boxplot() +
	labs(title = "Calories, by hotdog Type")

beef <- hotdogs[Type == "Beef"]
meat <- hotdogs[Type == "Meat"]
poultry <- hotdogs[Type == "Poultry"]

with(beef, t.test(Calories)$conf.int)
with(meat, t.test(Calories)$conf.int)
with(poultry, t.test(Calories)$conf.int)

t.test(beef$Calories, alternative = "less", conf.int = 0.99)
t.test(meat$Calories, alternative = "less", conf.int = 0.99)
t.test(poultry$Calories, alternative = "less", conf.int = 0.99)

```

## 2.)

> Find a 95% confidence interval for the variance in the amount of calories found for each type of hotdog:
beef, meat and poultry.


```{r data_2, echo=T}

var.conf.int = function(x, conf.level = 0.95) {
	df <- length(x) - 1
	chilower <- qchisq((1 - conf.level) / 2, df, lower.tail = TRUE)
	chiupper <- qchisq((1 - conf.level) / 2, df, lower.tail = FALSE)
	v <- var(x)
	c(df * v / chiupper, df * v / chilower)
}

pretty_vector(with(beef, var.conf.int(Calories)))
pretty_vector(with(meat, var.conf.int(Calories)))
pretty_vector(with(poultry, var.conf.int(Calories)))

```
