---
title: "Home Sales Analysis"
output: html_document
---

Source Data: <b>home_sales.csv</b>

```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(gplots, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

lp.w <- "E:/GitHub/R-Playground"
lp.h <- "C:/Projects/R/Playground"

if (file.exists(lp.w)) {
	base.dir <- lp.w
} else if (file.exists(lp.h)) {
	base.dir <- lp.h
}

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/_Data Files/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
  header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

home.sales <- loadDataFile("home_prices.csv")
str(home.sales)

```

<b>a.)</b> <i>What are the measurement levels of each of the eight variables ?</i>

```{r}

variable_types = data.frame(
      Variable = colnames(home.sales),
      Type = c("Ratio",
                "Ratio",
                "Ordinal",
                "Ordinal",
                "Ratio",
                "Nominal",
                "Nominal",
                "Ratio"))

print(variable_types)

```

<b>b.)</b> <i>Should any variable have its values changed to better reflect its true nature ?</i>

```{r}

home.sales$Year <- as.Date(paste(home.sales$Year, 12, 31, sep = "-"))
home.sales$Baths <- factor(home.sales$Baths, ordered = TRUE)
home.sales$Feats <- factor(home.sales$Feats, ordered = TRUE)
home.sales$NBR <- home.sales$NBR == "YES"
home.sales$Corner <- home.sales$Corner == "YES"

str(home.sales)

```

<b>c.)</b> <i>From the vector “price”, select a simple random sample of size 12. Assign the sample to the name “SRS.”"</i>

```{r}

set.seed(9999)
SRS <- sample(home.sales$Price, 12)

print(SRS)

mu_SRS <- mean(SRS)
print(mu_SRS)

```

<i>Print SRS and determine the mean value.</i>

```{r}

mu_SRS <- mean(SRS)

print(paste('Sample mean: ', round(mu_SRS,2)))

```

<b>d.)</b> <i>From the vector “price”, select a systematic sample of twelve observations. Start with the seventh
observation and pick every 10th observation thereafter (i.e. 7, 17, 27,). You should end with the 117th
observation. Assign the sample vector to “SS.”" Print the values of SS and compute the mean value.</i>

```{r}

obs <- seq(from = 7, to = 117, by = 10)
print("Systematic Sample: ")
print(obs)

SS <- home.sales[obs,]$Price

print(SS)

mu_SS <- mean(SS)
print(paste("SS Mean: ", round(mu_SS,2)))

```

e) Examine the printed values and mean values obtained from the two sampling procedures. Do you see a
difference? Try the commands summary(SRS) and summary(SS).

```{r}

print('SRS Summary')
summary(SRS)

print('SS Summary')
summary(SS)

```

<b>f.)</b> <i>Create histograms and stem-and-leaf plots for SRS and SS using hist() and stem().</i>

```{r, echo=FALSE}

stem(SRS)

pSRS <- qplot(SRS,
      geom = "histogram",
      bins = 5,
      main = "Histogram for Random Sample Price",
      xlab = "Price")

stem(SS)

pSS <- qplot(SS,
      geom = "histogram",
      bins = 5,
      main = "Histogram for Systematic Sample Price",
      xlab = "Price")

grid.arrange(pSRS, pSS, nrow = 1)

```

How do the two samples compare?